---
title: 'Introduction to R and Statlab Server'
short-title: "STAT 494"
author: "Xuelong Wang"
date: '`r format(Sys.Date(), "%B %d, %Y")`'      # Month DD, YYYY (Main Slide)
short-date: '`r format(Sys.Date(), "%m/%d/%Y")`' # MM/DD/YYYY (Lower Right)
section-titles: false                            # Provides slide headings
safe-columns: true   # Enables special latex macros for columns.
header-includes:
   - \usepackage{amsmath, bbm}
output: 
   uiucthemes::beamer_illinois: 
       toc: true
       fig_height: 6
---

```{r setup, include=FALSE}
options(width = 40)
library(doParallel)
library(foreach)
library(parallel)
library(iterators)
```

# Introduction to the Statlab Server

### How to connect to the Statlab Server

* Fixed IP address 131.193.178.77
* hostname: statlab.math.uic.edu 
* Mac users 
    + Terminal.app
    + ssh username@statlab.math.uic.edu
* Windows users 
    + PuTTY
    + open source SSH client for windows
    + http://www.chiark.greenend.org.uk/~sgtatham/putty/download.html
    
### How to use the R.Studio 

* R 
    + Using Terminal or PuTTY 
    + just type R
* Rstudio 
    + Web browser e.g. Chrome 
    + http://statlab.math.uic.edu:8787/
* some notes about using Rstudio 
    + save all the documents 
    + save and delete all objects in the Global Enviroment (rm(list=ls()))
    + using top and kill the terminate your R sessions

# Introduction to R with an example

### Basics concepts of R

#### Object-oriented

R is an object-oriented language

- Everything in R is an object and each object has a class e.g. matrix, list, data.frame,etc...

- For different classes, R has different operations which are called "Method"


### An example of Linear Regression

#### Mathematical equation

\[
  Y = \beta_0 + \beta_1X + \epsilon,
\] 
Where $\beta_0$ is the intercept and $\beta_1$ is the slope

#### Goal

Estimate the coefficients and prediction

### Example Cars' Speed and distances to stop 

```{r, eval=FALSE}
?cars
```

#### Mathematical equation

\[
  Y_{dist} = \beta_0 + \beta_{speed}X_{speed} + \epsilon,
\] 
Where $\beta_0$ is the intercept and $\beta_{speed}$ is the slope

#### Goal

Using a car's speed to predict the distance it will use to stop

- esitmate the $\beta_0$ and $\beta_{speed}$


### Check the data

```{r check data, collapse=TRUE}
head(cars)
class(cars)
```

### Plot X and Y together 

```{r plot x and y, collapse=TRUE}
plot(x = cars$speed, y = cars$dist)
```

### Plot a box-plot of speeds for outliers
```{r plot speed, collapse=TRUE}
boxplot(cars$speed, main="Speed")
```

### Plot a box-plot of distances for outliers
```{r plot distance, collapse=TRUE, fill = TRUE}
boxplot(cars$dist, main="Distance")
```

### Fit the model 

```{r fit the model, collapse=TRUE, tidy=TRUE}
help(lm)
class(lm)
linearMod <- lm(dist ~ speed, data=cars)
class(linearMod)
print(linearMod)
```

### plot the fitted model

```{r plot the model, collapse=TRUE, tidy=TRUE}
par(mfrow = c(2, 2))
plot(linearMod)
```

### Using the fitted model to predict

```{r prediction, collapse=TRUE, tidy=TRUE}
help(predict)
help(predict.lm)
# esimation
est <- predict(linearMod)
est[1:5]
# prediction
new_data <- data.frame(speed = 200)
pre <- predict(linearMod, newdata = new_data)
pre
```

### plot the predicted value vs true

```{r predicted vs true, collapse=TRUE, tidy=TRUE}
plot(x = cars$speed, y = cars$dist)
lines(x = cars$speed, linearMod$fitted.values, type='l', col='blue')
```

# Introduction to parallel computation 

### *foreach* package

#### An easy and starndard way of parallel comuptation

- Can run a for-loop task as a set of of parallel tasks 

- Take care of the communication between the tasks (cores)

### Getting start Example

Calculate the sum of the square $$\sum_{i=1}^{10000}\sum_{j=1}^{i} j^2$$
There is a warning saying the loop ran sequentially   
To run the loop parallelly,  we need to register parallel backends. 

```{r, tidy = TRUE, fill =TRUE, collapse=TRUE, tidy.opts=list(width.cutoff = 50)}
system.time(foreach(i=1:10000) %do% sum((1:i)^2))[3]
system.time(foreach(i=1:10000) %dopar% sum((1:i)^2))[3]
```

### Parallel backends

```{r, tidy = TRUE, fill =TRUE, collapse=TRUE}
registerDoParallel()
getDoParWorkers()
```

*registerDoParallel()* is used to register cores to parallel computation 

```{r, tidy = TRUE, fill =TRUE, collapse=TRUE, tidy.opts=list(width.cutoff = 50)}
system.time(foreach(i=1:10000) %do% sum(sqrt(1:i)))[3]
system.time(foreach(i=1:10000) %dopar% sum(sqrt(1:i)))[3]
```

### Bootstraping example
```{r, echo=FALSE, eval=TRUE}
x <- iris[which(iris[,5] != "setosa"), c(1,5)]
dopar <- system.time(r <- foreach(i = 1:10000) %dopar% {
  ind <- sample(100, 100, replace=TRUE)
  result1 <- glm(x[ind,2]~x[ind,1],
                 family=binomial(logit))
  coefficients(result1)
  })[3]
do <- system.time(r <- foreach(i = 1:10000) %do% {
  ind <- sample(100, 100, replace=TRUE)
  result1 <- glm(x[ind,2]~x[ind,1],
                 family=binomial(logit))
  coefficients(result1)
  })[3]
```

```{r, tidy = TRUE, fill =TRUE, collapse=TRUE, tidy.opts=list(width.cutoff = 45)}
dim(x)
r <- foreach(i = 1:10000) %dopar% {
  ind <- sample(100, 100, replace=TRUE)
  result1 <- glm(x[ind,2]~x[ind,1], family=binomial(logit))
  coefficients(result1)
  }
```
- Escaped time for using  `r getDoParWorkers()` cores is `r dopar` seconds
- Escaped time for using single core is `r do` seconds
